---
title: "dataset-description"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
```

## Data Loading and Processing

https://archive.ics.uci.edu/ml/datasets/Diabetes+130-US+hospitals+for+years+1999-2008?fbclid=IwAR07N48YplnMhoLF3NTeuFQm65xi97lmjX83pTnysFARSVqb-LdmIYCT3Ys

```{r}
castNum <- function(n, num.places) {
    x<-as.numeric(as.character(n))
    if (is.na(x)){
      return(n)
    }else{
      return(as.character(floor(x/10^num.places)*10^num.places))
    }
}
truncate.codes <- function(dataset, num.places) {
  dataset$diag_1 <- map2(dataset$diag_1, num.places, castNum) %>% unlist()
  dataset$diag_2 <- map2(dataset$diag_2, num.places, castNum) %>% unlist()
  dataset$diag_3 <- map2(dataset$diag_3, num.places, castNum) %>% unlist()
  dataset %>% mutate(
    diag_1 = ifelse(str_length(diag_1) == 1 & diag_1 != '?', str_c('00', diag_1), diag_1),
    diag_2 = ifelse(str_length(diag_2) == 1 & diag_2 != '?', str_c('00', diag_2), diag_2),
    diag_3 = ifelse(str_length(diag_3) == 1 & diag_3 != '?', str_c('00', diag_3), diag_3),
  ) %>% mutate(
    diag_1 = ifelse(str_length(diag_1) == 2 & diag_1 != '?', str_c('0', diag_1), diag_1),
    diag_2 = ifelse(str_length(diag_2) == 2 & diag_2 != '?', str_c('0', diag_2), diag_2),
    diag_3 = ifelse(str_length(diag_3) == 2 & diag_3 != '?', str_c('0', diag_3), diag_3),
  )
  write.csv(dataset, file=str_c('dataset_diabetes/diabetic_data_cleaned', num.places, '.csv'))
  dataset
}
dataset <- read.csv('dataset_diabetes/diabetic_data.csv')
icd.codes <- read.csv('icd_codes.csv')
diabetes.diagnoses <- c('249', '250', '648')
```

## Dataset Overview

```{r}
num.unique.encounters <- dataset %>% select(encounter_id) %>% unique() %>% count()
num.unique.patients <- dataset %>% select(patient_nbr) %>% unique() %>% count()
num.diabetic.patients <- dataset %>% select(patient_nbr, diag_1, diag_2, diag_3) %>% gather("diagnosis_num", "diagnosis", -patient_nbr) %>% group_by(patient_nbr) %>% summarise(any(diagnosis %in% diabetes.diagnoses)) %>% count()
num.medication.changed <- dataset %>% filter(change == 'Ch') %>% count()
num.readmitted <- dataset %>% filter(readmitted != "NO") %>% count()
```

The dataset contains `r num.unique.encounters` unique encounters over `r num.unique.patients` unique patients. There are `r num.diabetic.patients` diabetic patients in the dataset. Medications were changed in `r num.medication.changed` of the admissions. `r num.readmitted` admissions did not result in readmission.

### Primary Diagnoses

```{r warning=FALSE}
processed.data <- truncate.codes(dataset, 0) # process diagnosis codes to the nearest integer
primary.diagnoses <- processed.data %>% group_by(diag_1) %>% count() %>% arrange(desc(n)) %>% left_join(icd.codes, by = c('diag_1' = 'Code'))
primary.diagnoses %>% head(50) %>% kable()
```

### Number of Admissions of Patients Taking Diabetes Medication

```{r}
taking.diabetes.medications <- dataset %>% group_by(diabetesMed) %>% count()
taking.diabetes.medications %>% kable()
```

